---
title: "Alpha optimization formula, simulation and real data analysis"
date: 2019-10-08
output: 
  pdf_document:
    toc: true # table of content true
    toc_depth: 2
---


```{R include = FALSE}
setwd('/Users/yaolanqiu/Desktop/NYU/Research/FDA/alpha_optimization_2019-10-05/cluster/result')
library(knitr)
library(png)
library(grid)
```


# Purity calculation and alpha optimization 

More informaiton please see the other pdf file.

# Simulation 

## Data generation 


# Simulation Setting

#### Sample size

* 200 subjects in each group 

#### Angle between $\Gamma_1$ and $\Gamma_3$

$0, 30,  60, 90, 120, 150, 180$ degrees. 

#### Dimensions

p = 4

#### Parameters

* Each subject has 7 time points: $2 \times 200 \times 7 = 2800$  

* True $\alpha = [0.5, 0.5, 0.5, 0.5]_4$

* $\beta_{pbo} = [0, 3, 0.9]^{\prime}, \beta_{drg} = [0, 3.1, 1]^{\prime}$

* $\Gamma_{drg} = [0,1,0]^{\prime}$, angle between the two $\Gamma$ lines can be 0, 30, 60, 90, 120, 150, and 180 degrees. 

* $|\Gamma_{drg}| = |\Gamma_{pbo}| = 1$



## LME estimation 

```{R include = FALSE}

for(readindata in 1){
  setwd('/Users/yaolanqiu/Desktop/NYU/Research/FDA/Embarc/dataset')
  dat = read.csv('changeScore8wPerProtocol166nWith20Primary.csv')
  head(dat)
  dim(dat)
  dat_temp = dat #[,c('ProjectSpecificId', 'hamd17_baseline', 'dur_MDE')]
  setwd('/Users/yaolanqiu/Desktop/NYU/rotation/Rotation2/week1/first code')
  newdat = read.csv("longFormat_score17.csv", header=T)
  merged = merge(newdat,dat_temp,
                 by = c('ProjectSpecificId', 'site', 'Stage1TX', 'age_evaluation'))
  dim(merged)
  dim(newdat)
  merged$trt = (merged$Stage1TX=="SER/CIT")*1+1
  
  dat = merged
  dim(dat)
  dat = na.omit(dat)
  dim(dat)
  
  dat0 = dat
}

dat$tt = dat$week
dat$subj = as.character(dat$ProjectSpecificId)
dat$y = dat$score17

dat_try = dat

```


Renew: 

\begin{itemize}
\item Use Lagrange multiplier to calculate the purity with restriction. 
\item Fit one LME to make the $D_1 = D_2$
\end{itemize}

If we want to set two $\hat D$ to be the same in two groups to make the model more stable (less variance), we could fit the following LME model: 


$$Y_i =S (\beta_i + b_i + \Gamma(\alpha_i^{\prime}x_i))  + \epsilon = S (\beta_i + \Gamma(\alpha_i^{\prime}x_i)) + S b_i + \epsilon$$

$$\begin{aligned}
Y = & S (\beta_1 + \Gamma_1(\alpha^{\prime}x)) trt + S (\beta_2 + \Gamma_2(\alpha^{\prime}x)) + S b + \epsilon  \\
= & S \big(\beta_1 trt + \beta_2 + \Gamma_1(\alpha^{\prime}x) trt + \Gamma_2(\alpha^{\prime}x) \big) + S b + \epsilon  
\end{aligned}$$

That is, 

* $\beta_{drg} = \beta_1 + \beta_2$, $\beta_{pbo} = \beta_2$

* $\Gamma_{drg} = \Gamma_1 + \Gamma_2$, $\Gamma_{pbo} = \Gamma_2$

* $D = D_1 = D_2$

For example: 

```{R include = FALSE}
library(lme4)
library(knitr)
dat_try = dat

levels = c()
for(i in 1:dim(dat)[2]){
  levels = c(levels, length(unique(dat[,i])))
}

covar.set = colnames(dat)[levels > 5]
covar.set = covar.set[c(2,4, 6,7,8,9,10,11:21)]

i = 10; j = 13; k = 15
selected.covar = covar.set[c(i,j, k)]
p = length(selected.covar)

alpha = 1

alpha = matrix(alpha,p,1)
alpha = alpha/c(sqrt(t(alpha) %*% alpha))
alpha = round(alpha, 2)

# calculate W = alpha^T x
covar_list = dat_try[,selected.covar]
covar_list = scale(covar_list)

temp_W = unlist(c(covar_list))
temp_W = matrix(temp_W, dim(dat_try)[1], p)
dat_try$W = c(temp_W %*% alpha)
```

Fit the model with one D

```{R}
fit = lmer(y ~ tt + I(tt^2) +
               W + W * tt + W * I(tt^2) + 
               trt * tt + trt *  I(tt^2) +  
               trt * W * tt + trt * W * I(tt^2) + 
               (tt + I(tt^2)|subj),
               data = dat_try, REML = FALSE)
summary(fit)$coefficient
```

Results:

```{R}
fit_cov = as.matrix(fixef(fit))
    
    beta1 = fit_cov[c("(Intercept)", "tt", "I(tt^2)"),] +
      fit_cov[c("trt", "tt:trt", "I(tt^2):trt"),] * 2
    beta2 = fit_cov[c("(Intercept)", "tt", "I(tt^2)"),] + 
      fit_cov[c("trt", "tt:trt", "I(tt^2):trt"),]
    
    gamma1 = fit_cov[c("W","tt:W", "I(tt^2):W"),] + 
      fit_cov[c("W:trt", "tt:W:trt", "I(tt^2):W:trt"),] * 2
    gamma2 = fit_cov[c("W","tt:W", "I(tt^2):W"),] + 
      fit_cov[c("W:trt", "tt:W:trt", "I(tt^2):W:trt"),]
```

```{R}
  beta1
as.matrix(VarCorr(fit)$subj)[1:3, 1:3] 
```

Fit the model with two Ds


```{R}
  dat_pbo_est = dat_try[dat_try$trt == 1, ]
  dat_drg_est = dat_try[dat_try$trt == 2, ]
  
  fit_pbo_est = lmer(y ~ tt + I(tt^2) + W + W * tt +
                                  W * I(tt^2) + (tt+I(tt^2)|subj),
                                data = dat_pbo_est, REML = FALSE)
  fit_drg_est = lmer(y ~ tt + I(tt^2) + W + W * tt +
                                  W * I(tt^2) + (tt+I(tt^2)|subj),
                                data = dat_drg_est, REML = FALSE)
```

```{R}
  beta1 = as.matrix(fixef(fit_drg_est))[1:3]
  D1 = as.matrix(VarCorr(fit_drg_est)$subj)[1:3, 1:3] 
  D2 = as.matrix(VarCorr(fit_pbo_est)$subj)[1:3, 1:3] 
  beta1
  D1
  D2
```



## Result

```{R echo= FALSE}
setwd('/Users/yaolanqiu/Desktop/NYU/Research/FDA/alpha_optimization_2019-10-05/cluster/result')
load('result_table.RData')
kable(round(result_table,2))
```

Where 

* angles: the angle between $\Gamma_1$ and $\Gamma_2$

* cos: the cosine similiarity. 

* true_KL: the true purity 

* KL: the mean estimated purity with two $D$ matrix estimation 

* KL2: the mean estimated purity with one $D$ matrix estimation 

#### Histograms 

* Estimated purity with two $D$ matrix estimation 

```{r fig.width=8, fig.height=6,echo=FALSE,  fig.align = "center"}
setwd('/Users/yaolanqiu/Desktop/NYU/Research/FDA/alpha_optimization_2019-10-05/cluster/result')
img <- readPNG("hist_2D.png")
grid.raster(img)
```

* Estimated purity with one $D$ matrix estimation 

```{r fig.width = 8, fig.height=6 ,echo=FALSE,  fig.align = "center"}
setwd('/Users/yaolanqiu/Desktop/NYU/Research/FDA/alpha_optimization_2019-10-05/cluster/result')
img <- readPNG("hist_1D.png")
grid.raster(img)
```



### Purity with different number of covariaes. 


I also simulate a dataset with the same parameters settings, while 

* sample size for each group is 1000

* Angle between $\Gamma_1$ and $\Gamma_2$ is 60

* p = 10

With different number of covariates added in the model, will the purity have difference? Will the purity increase? 

With one covariate included:

```{R echo = FALSE}
setwd('/Users/yaolanqiu/Desktop/NYU/Research/FDA/alpha_optimization_2019-10-05/cluster/result')
load('multi.RData')
kable(multi$O_t)
```

With more covariates included:

```{R echo = FALSE}
setwd('/Users/yaolanqiu/Desktop/NYU/Research/FDA/alpha_optimization_2019-10-05/cluster/result')
load('multi.RData')
kable(multi$multi_table)
```


# Embarc data analysis 


## Covariates 



We would like to focus on  the continuous variables first. Therefore, the covarites with level larger than 5 are included. 

The covariates names are: 

* "age_evaluation",  "hamd17_baseline", "dur_MDE", "age_MDE",  "axis2", "anger_attack"  ,  "anxious"  

As well as the behavior covariates: 

|Covariate name| Description | 
|:----------|--------------------------------------------------:|
|w0_4165|	A not B Interference Reaction Time in negative trials|
| w0_4167|	A not B Interference Reaction Time in non-negative trials|
|w0_4163|	A not B Interference Reaction Time in all trials|
|w0_4162|	A not B Itotal number of correct trials|
|w0_4169|	Median Reaction time for correct trials in the Choice reaction time task |
|w0_1844|	Number of valid recalled words in the Word Fluency task|
|w0_1916|	Flanker Accuracy, an  Accuracy effect is a measure of interference effects; Higher scores are indicative of increased interference effects (i.e., reduced cognitive control). |
|w0_1915|	Flanker Reaction Time, a  measure of interference effects; Higher scores are indicative of increased interference effects (i.e., reduced cognitive control).|
|w0_1920|	Accuracy effect, it measures post-conflict behavioral adjustments; Higher values indicate better cognitive control|


## Purity with one covaraite

```{R echo = FALSE}
setwd('/Users/yaolanqiu/Desktop/NYU/Research/FDA/Embarc/Embarc-optim-2019-10-05')
load('alt_old_one_table2.RData')
one_table[,2:4] = round(one_table[,2:4] ,3)
colnames(one_table) = c('covariates', '1','Purity with 2D','Purity with 1D')
kable(one_table[,c(1,3,4)])
```


## Purity with two covaraites

Then randomly select two covariates, i.e, "w0_1916" and "w0_1915", the purity is: 

With two D, $D_1, D_2$

```{R echo = FALSE}
setwd('/Users/yaolanqiu/Desktop/NYU/Research/FDA/Embarc/Embarc-optim-2019-10-05')
load('two_table2.RData')
two_table$temp_res1$value
```

With one D

```{R echo = FALSE}
two_table$temp_res2$value
```

## Purity with selected three covaraites 


Then randomly select three covariates, i.e, i.e. "w0_4163" "w0_1844" "w0_1915", 
the purity is: 

With two D, $D_1, D_2$

```{R echo = FALSE}
setwd('/Users/yaolanqiu/Desktop/NYU/Research/FDA/Embarc/Embarc-optim-2019-10-05')
load('three_table2.RData')
three_table$temp_res1$value
```

With one D

```{R echo = FALSE}
three_table$temp_res2$value
```


## Purity with all covariates

With two D, $D_1, D_2$

```{R echo = FALSE}
load('all_table2.RData')

all_table$temp_res1$value
```

With one D
```{r echo = FALSE}
all_table$temp_res2$value
```

