---
title: "Simulation to estimating alpha 2"
date: 2019-09-11
fontsize: 12pt
output: pdf_document
geometry: tmargin=1.8cm,bmargin=1.8cm,lmargin=2.1cm,rmargin=2.1cm
---

# Kullback-Leibler divergence

For distributions $F_1$ and $F_2$ of a continuous random variable, the KL divergence is defined as: 
\begin{equation} \label{eq0}
D_{KL}(F_1 || F_2) = \int_{-\infty}^{+\infty}f_1(x)log(\frac{f_1(x)}{f_2(x)})dx
\end{equation}
where $f_1$ and $f_2$ denote the probability density of $F_1$ and $F_2$. 

Therefore, we defined the *puirty* as
$$\begin{aligned}
& \int log(f_1) f_1 - \int log(f_2) f_1 + \int log(f_2) f_2 - \int log(f_1) f_2 \\
& =\big(-\frac{p}{2} log(2\pi) - \frac{1}{2}log(|  D_1|) -\frac{p}{2}\big) \\
& -\big( -\frac{p}{2} log(2\pi) - \frac{1}{2}log(|  D_2|) -\frac{1}{2}\big( tr(  D_2^{-1}   D_1) + (  \mu_1 -  \mu_2)^{\prime}  D_2^{-1}(  \mu_1 -  \mu_2)\big)\big) \\
&+ \big(-\frac{p}{2} log(2\pi) - \frac{1}{2}log(|  D_2|) -\frac{p}{2}\big) \\
&- \big(-\frac{p}{2} log(2\pi) - \frac{1}{2}log(|  D_1|) -\frac{1}{2}\big( tr(  D_1^{-1}   D_2) + (  \mu_1 -  \mu_2)^{\prime}  D_1^{-1}(  \mu_1 -  \mu_2)\big)\big) \\
& = -p + \frac{1}{2} tr(  D_2^{-1}   D_1) + \frac{1}{2} tr(  D_1^{-1}   D_2) + \frac{1}{2} (  \mu_1 -  \mu_2)^{\prime} (  D_1^{-1} +   D_2^{-1}) (  \mu_1 -  \mu_2)
\end{aligned}$$
where $\mu_1 =  \beta_1 +   \Gamma_1   \alpha^{\prime}   x$, 
$\mu_2 =   \beta_2 +   \Gamma_2   \alpha^{\prime}   x$.

That is:
$$\text{KL purity } = -p + \frac{1}{2} tr(  D_2^{-1}   D_1) + \frac{1}{2} tr(  D_1^{-1}   D_2) + \frac{1}{2} (  \mu_1 -  \mu_2)^{\prime} (  D_1^{-1} +   D_2^{-1}) (  \mu_1 -  \mu_2)$$

The optimization of $\alpha$: 
$$\alpha  = -2B^{-1}A^{\prime}$$
where

* $A = \sum_i (  \Gamma_1 -   \Gamma_2) ^{\prime}(  D_1^{-1} +   D_2^{-1}) (  \beta_1 -   \beta_2 )    x_i ^{\prime}$

* $B = \sum_i   x_i  \big((  \Gamma_1 -   \Gamma_2)  \big) ^{\prime}(  D_1^{-1} +   D_2^{-1}) \big((  \Gamma_1 -   \Gamma_2) \big)  x_i ^{\prime}$




# Simulation Setting

Sample size:

* 200 subjects in each group, 400 in total. 

Dimensions:

* p = 4. (4 baseline biosignature)

Angle between $\Gamma_1$ and $\Gamma_3$: 

* scenario: $0, 30,  60, 90, 120, 150, 180$ degrees.

Initial value:  True $\alpha$

Simulation time: 500

Parameters

* Each subject has 7 time points: $2 \times 100 \times 7 = 1400$  

* True $\alpha = [0.5, 0.5, 0.5, 0.5]^{\prime}$

* $\beta_{pbo} = [0, 3, 0.9]^{\prime}, \beta_{drg} = [0, 3.1, 1]^{\prime}$

* $\Gamma_{drg} = [0,1,0]^{\prime}$, angle between the two $\Gamma$ lines can be 0, 30, 60, 90, 120, 150, and 180 degrees. 

* $|\Gamma_{drg}| = |\Gamma_{pbo}| = 1$


# Result

```{R include = FALSE}
library(knitr)
library(png)
library(grid)
```

#### $\alpha$ calculated by $\hat \beta_{drg}, \hat \beta_{pbo}, \hat \Gamma_{drg}, \hat \Gamma_{pbo}, \hat D_{drg}, \hat D_{pbo}$

```{R echo = FALSE}
setwd('/Users/yaolanqiu/Desktop/NYU/Research/FDA/alpha_optimization_2019-09-07/cluster2/FDA_alpha2')
load('res_d.RData')
colnames(res_d) = c('Angle','Mean KL dif','Mean KL', 'SD KL', 'Mean cosine', 'SD cosine')
kable(res_d)
```

* Mean_KL_dif: the mean absolute differences between true *purity*  and the estimated $\hat {purity}$ during the 500 times replication. 

* Mean_KL: The mean value of the estimated *purity* 

* SD_KL: The standard deviation value of the estimated  *purity*

* Mean_cosine: the mean value of the cosine_similarity. 

* SD_cosine: The standard deviation value of the cosine_similarity


##### The histogram of purity

```{r fig.width=8, fig.height=4,echo=FALSE,  fig.align = "center"}
img <- readPNG('/Users/yaolanqiu/Desktop/NYU/Research/FDA/alpha_optimization_2019-09-07/cluster2/FDA_alpha2/KL.png')
grid.raster(img)
```


##### The histogram of the consine similarity 


```{r fig.width=8, fig.height=4,echo=FALSE,  fig.align = "center"}
img <- readPNG('/Users/yaolanqiu/Desktop/NYU/Research/FDA/alpha_optimization_2019-09-07/cluster2/FDA_alpha2/D.png')
grid.raster(img)
```


#### $\alpha$ calculated by $\hat \beta_{drg}, \hat \beta_{pbo}, \hat \Gamma_{drg}, \hat \Gamma_{pbo},$ true $D_{drg},D_{pbo}$

```{R echo = FALSE}
setwd('/Users/yaolanqiu/Desktop/NYU/Research/FDA/alpha_optimization_2019-09-07/cluster2/FDA_alpha2')
load('res_d2.RData')
colnames(res_d2) = c('Angle','Mean KL dif','Mean KL', 'SD KL', 'Mean cosine', 'SD cosine')
kable(res_d2)
```


* Mean_KL_dif: the mean absolute differences between true *purity*  and the estimated $\hat {purity}$ during the 500 times replication. 

* Mean_KL: The mean value of the estimated *purity* 

* SD_KL: The standard deviation value of the estimated  *purity*

* Mean_cosine: the mean value of the cosine_similarity. 

* SD_cosine: The standard deviation value of the cosine_similarity


##### The histogram of purity

```{r fig.width=8, fig.height=4,echo=FALSE,  fig.align = "center"}
img <- readPNG('/Users/yaolanqiu/Desktop/NYU/Research/FDA/alpha_optimization_2019-09-07/cluster2/FDA_alpha2/trueKL.png')
grid.raster(img)
```


##### The histogram of the consine similarity 

```{r fig.width=8, fig.height=4,echo=FALSE,  fig.align = "center"}
img <- readPNG('/Users/yaolanqiu/Desktop/NYU/Research/FDA/alpha_optimization_2019-09-07/cluster2/FDA_alpha2/trueD.png')
grid.raster(img)
```

